---
title: Finding new wedding bops with {tidyclust} and {spotifyr}
author: ''
date: '2022-08-20'
slug: []
categories: 
- rstats
- tidymodels
tags: []
subtitle: ''
summary: ''
authors: []
lastmod: '2022-08-20T16:20:44-05:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  dpi = 500,
  fig.width = 9,
  fig.height = 6
)

# setup themes
extrafont::loadfonts(device = "win")
ggplot2::theme_set(
  ggplot2::theme_minimal(base_family = "Roboto Slab", 
                         base_size = 14) +
    ggplot2::theme(plot.title.position = "plot",
                   plot.background = ggplot2::element_rect(fill = "white", color = "white"),
                   plot.title = ggtext::element_markdown(),
                   plot.subtitle = ggtext::element_markdown())
)
```

Last November, I (finally) popped the big question and [proposed](https://www.instagram.com/p/Cc6ReimO2tH/?igshid=YmMyMTA2M2Y=)! Since then, my fiance and I have been diligently planning our wedding. While we have most of the big-ticket items checked off (venue, catering, photography, etc.), one area we still have more work to do is on the wedding playlist. We've [started putting together a playlist on spotify](https://open.spotify.com/playlist/66saUDfW5ggYD6JDAHbxyV), but it feels like it's come to a bit of a stand-still. Currently, there's a mix of zesty bops and tame songs on the playlist (we need to accommodate both our college friends and our grandparents!), but spotify's track recommender only wants to suggest tamer songs right now. Our goal is to have a full dance floor the entire night --- to achieve this, we can use [spotifyr](https://www.rcharlie.com/spotifyr/index.html) and the new [tidyclust](https://emilhvitfeldt.github.io/tidyclust/index.html) package to pull in the current playlist, cluster the songs based on their features, and find new songs based on the bop cluster.

```{r}
library(tidymodels)
library(tidyclust)
library(spotifyr)
```

If you'd like to follow along, I'd recommend installing the development versions of [parsnip](https://github.com/tidymodels/parsnip) and [workflows](https://github.com/tidymodels/workflows), as some of the functionality that interacts with tidyclust [isn't yet on CRAN](https://emilhvitfeldt.github.io/tidyclust/articles/k_means.html#setup). 

## Pulling in the playlist

[spotifyr](https://www.rcharlie.com/spotifyr/index.html) is an R interface to spotify's web API and gives access to a host of track features (you can follow [this tutorial](https://msmith7161.github.io/what-is-speechiness/) to get it setup). I'll use the functions `get_user_playlists()` and `get_playlist_tracks()` to pull in songs that are currently on our wedding playlist (appropriately named "Ding dong").

```{r}
# get the songs that are currently on the wedding playlist
ding_dong <- 
  get_user_playlists("12130039175") %>%
  filter(name == "Ding dong") %>%
  pull(id) %>%
  get_playlist_tracks() %>% 
  as_tibble() %>%
  select(track.id, track.name, track.popularity) %>%
  rename_with(~stringr::str_replace(.x, "\\.", "_"))
```
```{r, echo=FALSE}
ding_dong %>%
  slice_head(n = 10) %>%
  knitr::kable()
```

Spotify estimates quite a few [features](https://developer.spotify.com/documentation/web-api/reference/#/operations/get-several-audio-features) for each song in their catalog: speechiness (the presence of words on a track), acousticness (whether or not a song includes acoustic instruments), liveness (estimates whether or not the track is live or studio-recorded), etc. We can use `get_track_audio_features()` to get the features for each song based on its `track_id`. 
 
```{r}
# pull in track features of songs on the playlist
track_features <- 
  ding_dong %>%
  pull(track_id) %>%
  get_track_audio_features()

# join together
ding_dong <- 
  ding_dong %>%
  left_join(track_features, by = c("track_id" = "id"))
```

In my case, I'm interested in the energy and valence (positivity) of each song, so I'll select these variables to use in the cluster analysis.

```{r, echo=FALSE}
ding_dong %>%
  select(track_name, valence, energy) %>%
  slice_head(n = 10) %>%
  knitr::kable()
```

## Clustering with `tidyclust` 

Currently, the playlist covers a wide spectrum of of songs. For new songs on the playlist, I'm really just interested in songs similar to others in the top right corner of the below chart with high energy and valence.

```{r, echo=FALSE}
# how are valence/energy related?
obj <- 
  ding_dong %>%
  ggplot(aes(x = valence,
             y = energy,
             tooltip = track_name)) + 
  ggiraph::geom_point_interactive(size = 3.5, alpha = 0.5) +
  scale_x_continuous(labels = scales::label_percent(accuracy = 1)) +
  scale_y_continuous(labels = scales::label_percent(accuracy = 1)) +
  labs(title = "The current wedding playlist",
       subtitle = "Hover over each point to see the song's name!")

ggiraph::girafe(
  ggobj = obj,
  options = list(
    ggiraph::opts_tooltip(opacity = 0.8,
                          css = "background-color:gray;color:white;padding:2px;border-radius:2px;font-family:Roboto Slab;"),
    ggiraph::opts_hover(css = "fill:#1279BF;stroke:#1279BF;cursor:pointer;")
  )
)
```

Broadly, there are three generic categories that the songs on the current playlist fall into: high energy and valence, low energy, or low valence (songs with low energy and valence will fall into one of the "low" categories). Rather than manually assign categories, we can use tidyclust to cluster the songs into three groups using the [kmeans](https://en.wikipedia.org/wiki/K-means_clustering) algorithm. 

There's some [great documentation on the package site](https://emilhvitfeldt.github.io/tidyclust/articles/k_means.html), but to get started, we'll categorize the songs on the current playlist by "fitting" a kmeans model (using the [stats](https://stat.ethz.ch/R-manual/R-devel/library/stats/html/00Index.html) engine under the hood).

```{r}
# create a clustering obj
set.seed(918)
ding_dong_clusters <- 
  k_means(num_clusters = 3) %>%
  fit(~ valence + energy,
      data = ding_dong) 
```

```{r, echo=FALSE}
pal <- MetBrewer::MetPalettes$Egypt[[1]]

obj <- 
  ding_dong_clusters %>%
  augment(ding_dong) %>%
  ggplot(aes(x = valence,
             y = energy,
             color = .pred_cluster,
             tooltip = track_name)) +
  ggiraph::geom_point_interactive(size = 3.5, alpha = 0.75) +
  scale_x_continuous(labels = scales::label_percent(accuracy = 1)) +
  scale_y_continuous(labels = scales::label_percent(accuarcy = 1)) +
  theme(legend.position = "none") +
  labs(title = "Clusters in the current playlist",
       subtitle = glue::glue("Clustered into",
                             "**{riekelib::color_text('zesty bops',pal[1])}**,",
                             "**{riekelib::color_text('dark bangers', pal[3])}**,",
                             "and",
                             "**{riekelib::color_text('mellow jams', pal[2])}**",
                             .sep = " ")) +
  MetBrewer::scale_color_met_d("Egypt")

ggiraph::girafe(
  ggobj = obj,
  options = list(
    ggiraph::opts_tooltip(opacity = 0.8,
                          use_fill = TRUE,
                          css = "color:white;padding:2px;border-radius:2px;font-family:Roboto Slab;"),
    ggiraph::opts_hover(css = "fill:#1279BF;stroke:#1279BF;cursor:pointer;")
  )
)
```

As expected, the majority of songs in the current playlist fall into the bop cluster. Let's explore this cluster using in more detail with the custom metric `vibe`.

```{r}
# assign to clusters
ding_dong_vibes <- 
  ding_dong_clusters %>%
  augment(ding_dong) %>%
  select(track_name,
         valence, 
         energy, 
         .pred_cluster) %>%
  mutate(vibe = valence + energy)

# what are songs with the biggest vibe?
ding_dong_vibes %>%
  arrange(desc(vibe)) %>%
  slice_head(n = 10) %>%
  knitr::kable()
```

As expected, when arranging by `vibe`, the top songs are all a part of the first cluster. And they are, indeed, *a vibe*:

```{r, echo=FALSE}
blogdown::shortcode_html("spotify", type = "track", id = "1uPrIHgYztXSkkcts9jet8")
```

Compare that with the second cluster, which are generally lower energy (I'd personally disagree with spotify ranking Mr. Blue Sky and Single Ladies as "low energy," but most others make sense).

```{r}
ding_dong_vibes %>%
  filter(.pred_cluster == "Cluster_2") %>%
  arrange(vibe) %>%
  slice_head(n = 10) %>%
  knitr::kable()
```

```{r, echo=FALSE}
blogdown::shortcode_html("spotify", type = "track", id = "12jjuxN1gxlm29cqL5M6MW")
```

Finally, the third cluster mostly contains songs with low valence but relatively high energy.

```{r}
ding_dong_vibes %>%
  filter(.pred_cluster == "Cluster_3") %>%
  arrange(vibe) %>%
  slice_head(n = 10) %>%
  knitr::kable()
```

```{r, echo=FALSE}
blogdown::shortcode_html("spotify", type = "track", id = "3n3Ppam7vgaVa1iaRUc9Lp")
```
  
Now that I have the songs in the current playlist sorted by cluster, let's pull in some new songs and assign them to the appropriate cluster!

## Adding new songs

To go searching for new songs, we'll start by casting a wide net then narrow the search with some of the `get_*()` functions from spotifyr. I'll start by using `get_categories()` to explore the categories available in spotify.

```{r}
get_categories() %>%
  as_tibble() %>%
  select(id, name) %>%
  slice_head(n = 10) %>%
  knitr::kable()
```

I don't really want to play country music or R&B during the wedding, so I'll filter to a few categories before using `get_category_playlists()` to pull in the featured playlists available in each category. 

```{r}
# pull in playlist ids
playlists <- 
  get_categories() %>%
  as_tibble() %>%
  filter(id %in% c("toplists", "hiphop", "pop", "rock", "summer")) %>%
  pull(id) %>%
  map_dfr(get_category_playlists) %>%
  as_tibble() %>%
  select(id, name, description) %>%
  distinct(id, .keep_all = TRUE)

playlists %>%
  slice_head(n = 10) %>%
  knitr::kable()
```

There's a lot of playlists in `playlists`, so I've gone through and selected a few that I'm interested in exploring further.

```{r}
selected_playlists <-
  c("Today's Top Hits",
    "mint",
    "Top 50 - US",
    "Top 50 - Global",
    "Viral 50 - US",
    "Viral 50 - Global",
    "New Music Friday",
    "Most Necessary",
    "Internet People",
    "Gold School",
    "Hot Hits USA",
    "Pop Rising",
    "teen beats",
    "big on the internet",
    "Party Hits",
    "Mega Hit Mix",
    "Pumped Pop",
    "Hit Rewind",
    "The Ultimate Hit Mix",
    "00s Rock Anthems",
    "Summer Hits",
    "Barack Obama's Summer 2022 Playlist",
    "Summer Hits of the 10s",
    "Family Road Trip")
```

With this shorter list of playlists, I can pull in the all the songs that appear on each with `get_playlist_tracks()`. Some songs may appear on multiple playlists, so we'll only look at unique songs by `track_id`. I've already pulled in features for songs currently on the playlist, so we can filter those out as well. Finally, `get_track_audio_features()` limits queries to a maximum of 100 songs, so we'll select the top 100 most popular songs within the sample.

```{r}
new_songs <- 
  playlists %>%
  filter(name %in% selected_playlists) %>%
  pull(id) %>%
  map_dfr(get_playlist_tracks) %>%
  as_tibble()

new_songs <- 
  new_songs %>%
  select(track.id,
         track.name,
         track.popularity) %>%
  rename_with(~stringr::str_replace(.x, "\\.", "_")) %>%
  distinct(track_id, .keep_all = TRUE) %>%
  arrange(desc(track_popularity)) %>%
  filter(!track_id %in% ding_dong$track_id) %>%
  slice_head(n = 100)
```

```{r, echo=FALSE}
new_songs %>%
  slice_head(n = 10) %>%
  knitr::kable()
```

Now let's assign these 100 news songs to the clusters we found earlier based on their valence and energy!

```{r}
new_song_features <- 
  new_songs %>%
  pull(track_id) %>%
  get_track_audio_features()

new_songs <- 
  new_songs %>%
  left_join(new_song_features, by = c("track_id" = "id"))

new_songs_clustered <- 
  ding_dong_clusters %>%
  augment(new_songs) %>%
  select(track_name,
         valence,
         energy,
         .pred_cluster) %>%
  mutate(vibe = valence + energy)
```

```{r, echo=FALSE}
obj <- 
  new_songs_clustered %>%
  ggplot(aes(x = valence,
             y = energy,
             color = .pred_cluster,
             tooltip = track_name)) +
  ggiraph::geom_point_interactive(size = 3.5, alpha = 0.75) +
  scale_x_continuous(labels = scales::label_percent(accuracy = 1)) +
  scale_y_continuous(labels = scales::label_percent(accuarcy = 1)) + 
  theme(legend.position = "none") +
  labs(title = "New songs, same old clusters",
       subtitle = glue::glue("Clustered into",
                             "**{riekelib::color_text('zesty bops',pal[1])}**,",
                             "**{riekelib::color_text('dark bangers', pal[3])}**,",
                             "and",
                             "**{riekelib::color_text('mellow jams', pal[2])}**",
                             .sep = " ")) +
  MetBrewer::scale_color_met_d("Egypt")

ggiraph::girafe(
  ggobj = obj,
  options = list(
    ggiraph::opts_tooltip(opacity = 0.8,
                          use_fill = TRUE,
                          css = "color:white;padding:2px;border-radius:2px;font-family:Roboto Slab;"),
    ggiraph::opts_hover(css = "fill:#1279BF;stroke:#1279BF;cursor:pointer;")
  )
)
```

Nice! It looks like the new songs are far more broad than the original playlist, but we can look at just the songs in the first cluster with the biggest vibe.

```{r}
new_songs_clustered %>%
  filter(.pred_cluster == "Cluster_1") %>%
  arrange(desc(vibe)) %>%
  slice_head(n = 10) %>%
  knitr::kable()
```

Now for the true vibe check --- do these songs *belong* on the playlist?

```{r, echo=FALSE}
blogdown::shortcode_html("spotify", type = "track", id = "0xzI1KAr0Yd9tv8jlIk3sn")
```

Oh ***hell* yeah!**

>**Some notes**: this analysis was done on Aug. 8th, 2022 --- spotify's featured playlists and tracks change on on a regular basis and may also depend on unique user data. 
